**RR - PEER ASSIGNMENT 1**
========================================================
## A. DATA INFO
Dataset (from [RR webpage][ref]): activity monitoring, containing the variables:  
    * steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)  
    * date: The date on which the measurement was taken in YYYY-MM-DD format  
    * interval: Identifier for the 5-minute interval in which measurement was taken

[ref]:https://class.coursera.org/repdata-003/human_grading/view/courses/972142/assessments/3/submissions

```{r}
data<-read.csv("~/Desktop/Coursera-RR/Exercises/Assignment1/activity.csv")
str(data)
```

- R is reading variable *date* as factor; it should read as Date:
```{r}
data$date<-as.Date(data$date, format="%Y-%m-%d")
str(data)
# ok!
```

## B. NUMBER OF STEPS PER DAY

For this task, first of all, I can reorganize the dataset as to show the total number of steps per day:

```{r}
# Vector - total number of steps:
total<-tapply(data$steps, data$date, function(x)sum(x))
head(total)
totalsteps<-array(NA, c(length(total),1))
for(i in 1:length(total)){
    totalsteps[i]<-total[[i]]
}
head(totalsteps)

# Vector - days:
date_steps<-c(unique(data$date))
head(date_steps)

# Double-check type = Date:
str(date_steps)
# ok!
```

```{r}
# Create new data frame:
totalsteps_frame<-data.frame(date_steps,totalsteps)
str(totalsteps_frame)
```
- For the plot, I will use the package 'ggplot2':
```{r}
require(ggplot2)
```

**B.1. Histogram; total # steps per day**
```{r, fig.height=5, fig.width=7, message=FALSE}
ggplot(aes(x=totalsteps), data=totalsteps_frame)+
    geom_histogram(color="black", fill="white") + ylab("Frequency") + xlab("Total number of steps per day")
```

**B.2. Mean and median; total # steps per day**
- Mean:
```{r}
mean(totalsteps_frame$totalsteps, na.rm=T)
```
The mean is `r mean(totalsteps_frame$totalsteps, na.rm=T)`.

- Median:
```{r}
median(totalsteps_frame$totalsteps, na.rm=T)
```
The median is `r median(totalsteps_frame$totalsteps, na.rm=T)`.

## C. AVERAGE DAILY ACTIVITY PATTERN
For this task, I need the average of each first entry (i.e. first 5 min interval) of every day, the average of each second entry (i.e. second 5 min interval), etc.

```{r}
# Get average per interval:
averages<-tapply(data$steps, data$interval, function(x)mean(x, na.rm=T))
meansteps_interval<-array(NA, c(length(averages),1))
for(i in 1:length(averages)){
    meansteps_interval[i]<-averages[[i]]
}
head(meansteps_interval)

# Vector - interval:
interval<-unique(data$interval)
str(interval)

# Create new dataset:
interval_frame<-data.frame(meansteps_interval, interval)
str(interval_frame)
```

**C.1. Time series plot; average number of steps of each 5 min interval**
```{r, fig.height=5, fig.width=7}
ggplot(aes(x=interval, y=meansteps_interval), data=interval_frame)+
    geom_line() + xlab("5 min interval")+ylab("Average number of steps")
```

**C.2. Which 5 min interval has the highest average # of steps?**
```{r}
(d<-interval_frame$interval[max(interval_frame$meansteps_interval)])
```
The interval `r d` is the one with the highest average number of steps.

## D. IMPUTING NAs

**D.1. Total number of NAs**
```{r}
missing<-which(is.na(data$steps)==T)
length(missing)
```
The total number of missing values in the dataset is `r length(missing)`.

**D.2. Fill in NAs with mean values**
```{r}
# New vector:
steps2<-data$steps
steps2[is.na(steps2)==T]<-mean(data$steps, na.rm=T)
str(steps2)
```

**D.3. Create data set with ex. 2**
```{r}
data2<-data
data2$steps<-steps2
str(data2)
```

**D.4. Histogram total number of steps; report mean and median total number of steps/day; do these values differ from first part of the assignment (i.e. with NAs)? What is the impact of NAs?**  
As before, I need a new dataset indicating the total number of steps; this time, however, I will start off with *data2*, the dataset in which NAs were filled in with the means.

```{r}
# Vector - total number of steps:
total2<-tapply(data2$steps, data2$date, function(x)sum(x))
totalsteps2<-array(NA, c(length(total2),1))
for(i in 1:length(total2)){
    totalsteps2[i]<-total2[[i]]
}
head(totalsteps2)

# Vector - days = date_steps from section B
# Create new dataset
totalsteps_frame2<-data.frame(totalsteps2, date_steps)
str(totalsteps_frame2)
```

- Histogram:
```{r, message=FALSE}
ggplot(aes(x=totalsteps2), data=totalsteps_frame2)+
    geom_histogram(color="black", fill="white") + ylab("Frequency") + xlab("Total number of steps per day, no NAs")
```

- Mean:
```{r}
mean(totalsteps_frame2$totalsteps2)
```
The mean is `r mean(totalsteps_frame2$totalsteps2)`, and it is not differente than the mean obtained in section B. This is not surprising; considering how mean is calculated, if mean values are added to the data set, the mean should not change.

- Median:
```{r}
median(totalsteps_frame2$totalsteps2)
```
The median is `r median(totalsteps_frame2$totalsteps2)`, which is slightly different than the original median obtained in section B (i.e. `r median(totalsteps_frame$totalsteps, na.rm=T)`). Considering that the original mean (`r mean(totalsteps_frame$totalsteps, na.rm=T)`) and median were similar, the inclusion of mean values in the data set would not be expected to considerably affect the median, as observed. 

## E. WEEKDAYS vs. WEEKENDS
(Use filled-in NAs data set -> data2)

**1. Create variable with 2 factors: weekday and weekend**
```{r}
# Find days of the week:
data2$whatdayistoday<-weekdays(data2$date)
str(data2)

# Create variable:
data2$day<-data2$whatdayistoday
data2$day[data2$day=="Saturday" | data2$day=="Sunday"]<-"weekend"
data2$day[data2$day != "weekend"]<-"weekday"
str(data2)

# ok, but change to factors:
data2$day<-as.factor(data2$day)
```

**2. Time series plot of the average number of steps per interval in weekdays & weekends**  
First, I need a data set with the averages:
```{r}
# Vector - mean weekdays:
wds<-tapply(data2$steps[data2$day=="weekday"], data2$interval[data2$day=="weekday"], function(x)mean(x))
mean_weekdays<-array(NA, c(length(wds),1))
for(i in 1:length(wds)){
    mean_weekdays[i]<-wds[[i]]
}
head(mean_weekdays)

# Vector - mean weekends:
wks<-tapply(data2$steps[data2$day=="weekend"], data2$interval[data2$day=="weekend"], function(x)mean(x))
mean_weekends<-array(NA, c(length(wks),1))
for(i in 1:length(wks)){
	mean_weekends[i]<-wks[[i]]
}
head(mean_weekends)

# Vector - interval from before = interval
# Dataset
days_interval<-data.frame(interval, mean_weekdays, mean_weekends)
str(days_interval)
```

- For multiple plots in a single window with ggplot, I need to use function multiplot (available from [cookbook-r][ref]).
[ref]: http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)

```{r}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

- Time series for weekdays (top) and weekends (bottom):
```{r}
plot1<-ggplot(aes(x=interval, y=scale(mean_weekdays)), data=days_interval)+
    geom_line() + xlab("5 min interval") + ylab("Mean number of steps on weekdays") + theme_bw()
plot2<-ggplot(aes(x=interval, y=scale(mean_weekends)), data=days_interval)+
	geom_line() + xlab("5 min interval") + ylab("Mean number of steps on weekends") + theme_bw()
multiplot(plot1, plot2)
```
